The illusory promise of the Aligned Rank Transform

### 2025-12-20 text/50-ExperimentalMethod.qmd
and some changes

**Linear component.** The base of our data generation process for all distributions is the following linear term:

@@

**Generating ratio data.** We adopt generalized linear modeling [@Bolker2009glm] for ratio data, where we use an inverse link function $g^{-1}$ to make the connection between $\ell_{ijk}$ and the location parameter of the distribution of interest:  
$$
Y_{ijk} \sim \mathcal{F}(location=g^{-1}(\ell_{ijk}), ...)
$${#eq-glm-ratio}

where $\mathcal{F}$ denotes the probability density (or mass) function from which we draw the responses $Y_{ijk}$. For example, the mean of the exponential distribution is equal to $1 / \lambda$, where $\lambda$ is its rate parameter. This mean is linked to the linear term through the function $g^{-1}(x) = e^{x}$, and therefore, $\lambda = e^{-\ell_{ijk}}$. Or the link function commonly used for the binomial distribution is the *logit* function, but its inverse if the *logistic* function $g^{-1}(x) = 1 / (1 + e^{-x})$.

 @tbl-distributions presents how the location parameters of the other distributions are linked to the linear term. We emphasize that the Cauchy distribution does not define a GLM (since it does not belong to the exponential family), but we use a model that connects its location (median) through the identity link.

@@

: Distribution parameters and their links to the linear term $\ell_{ijk}$ {#tbl-distributions}

>>>>

**Linear component.** The base of our data generation process for all distributions is the following linear predictor:

@@

**Generating ratio data.** We generate ratio-scaled data using generalized linear modeling [@Bolker2009glm]. Specifically, we use an inverse link function $g^{-1}$ to relate the linear predictor $\ell_{ijk}$ to the location parameter of the target distribution:  
$$
Y_{ijk} \sim \mathcal{F}(location=g^{-1}(\ell_{ijk}), ...)
$${#eq-glm-ratio}

where $\mathcal{F}$ denotes the probability density (or mass) function from which the responses $Y_{ijk}$ are drawn. For example, the mean of the exponential distribution is equal to $1 / \lambda$, where $\lambda$ is its rate parameter. This mean is linked to the linear predictor through the inverse link function $g^{-1}(x) = e^{x}$, which implies $\lambda = e^{-\ell_{ijk}}$. Similarly, for the binomial distribution, the canonical link function is the *logit* function, whose inverse is the *logistic* function $g^{-1}(x) = 1 / (1 + e^{-x})$.

 @tbl-distributions summarizes how the location parameters of all the distributions are linked to the linear predictor. We emphasize that the Cauchy distribution does not define a generalized linear model  --- because it does not belong to the exponential family --- but we nevertheless use a model that links its location (the median) to the linear predictor via the identity link.

@@

: Distribution parameters and their links to the linear predictor $\ell_{ijk}$ {#tbl-distributions}
